Massively Parallel Monte Carlo Methods for Discrete Linear and Nonlinear Systems

### 2012-10-31 doc/prelim/presentation/prelim_presentation.tex
working through fanm slides

@@

@@

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{The FANM Method}

@@

>>>>

  \begin{itemize}
  \item A form of inexact Newton methods
  \end{itemize}

  \[
  ||\ve{J}(\ve{u}^k)\delta \ve{u}^k + \ve{F}(\ve{u}^k)|| \leq \eta^k
  ||\ve{F}(\ve{u}^k)||,\ \eta^k \in [0,1)
  \]

  \begin{itemize}
  \item Choose a Krylov method to solve for the Newton correction
  \item GMRES with a long recurrence relation obeserved as more robust
  \item Generates a monotonically decreasing residual from maintaining
    the optimization and orthogonality conditions
  \end{itemize}

  \pause
  Where does the Jacobian come from?
  \pause
  \begin{itemize}
  \item The Jacobian can come from hand-coded derivatives
    \begin{itemize}
    \item Tedious and error prone
    \item Repeated for each equation set and hard to do for multiphysics
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Krylov methods only need the action of the linear operator
  \item The action of the Jacobian can be approximated using a forward
    difference
  \end{itemize}

  \[
  \ve{J}(\ve{u})\ve{v} = \frac{\ve{F}(\ve{u} + \epsilon \ve{v}) -
    \ve{F}(\ve{u})}{\epsilon}
  \]

  \begin{itemize}
  \item Forms the basis of Jacobian-Free Newton-Krylov (JFNK) methods
    \begin{itemize}
    \item Sensitive to scaling and discretization error
    \item Still have to form part of Jacobian periodically for
      preconditioning
    \item Eventually break even with generating and storing the full
      Jacobian
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Automatically generate Jacobians from nonlinear function
    evaluations
    \begin{itemize}
    \item Overload math operators and apply the chain rule (FAD)
    \item Yields evaluations as accurate as function discretization
    \end{itemize}
  \item Modern packages take an element-level assembly approach
    \begin{itemize}
    \item Function evaluations are local, communication builds global
      data
    \end{itemize}
  \end{itemize}

  \[
  \ve{J}(\ve{u}) = \sum_{i=1}^N \ve{Q}^T_i \ve{J}_k \ve{P}_i
  \]
  \[
  e_k : \mathbb{R}^{n_k} \rightarrow \mathbb{R}^{m_k},\ \ve{J}_{k_i} =
  \partial e_{k_i} / \partial P_i u,\ \ve{P} \in \mathbb{R}^{m_{k_i}
    \times N}
  \]

  \begin{itemize}
  \item Performance studies give acceptable results for use in
    large-scale, production physics codes
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}[fragile]{The FANM Method}

  Forward-Automated Newton-MCSA

  \pause
  \begin{algorithm}[H]
    \begin{algorithmic}[1]
      \STATE $k := 0$
      \WHILE{$||\ve{F}(\ve{u}^{k})|| > \epsilon ||\ve{F}(\ve{u}^{0})||$}
      \STATE $\ve{J}(\ve{u}^{k}) \leftarrow AD(\ve{F}(\ve{u}^k))$
      \COMMENT{Automatic differentation}
      \STATE $\ve{J}(\ve{u}^k) \delta \ve{u}^k = -\ve{F}(\ve{u}^{k})$
      \COMMENT{Solve for the Newton correction with MCSA}
      \STATE $\ve{u}^{k+1} \leftarrow \ve{u}^k + \delta \ve{u}^k$
      \STATE $k \leftarrow k+1$
      \ENDWHILE
    \end{algorithmic}
    \caption{FANM}
  \end{algorithm}

  \pause
  \begin{itemize}
  \item Robustness of Newton's method (inexact)
  \item Accuracy and convenience of FAD
  \item Parallelism, memory, and resiliency benefits of MCSA
  \item Requires only nonlinear function evaluations
  \end{itemize}

@@

    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Jacobian will be in a compressed row storage format
      \end{itemize}
    \end{column}

    \begin{column}{0.5\textwidth}
      \small{\[ \ve{A} =
        \begin{bmatrix}
          2 & 0 & 8 & 0 & 0 & 0 \\ 4 & 5 & 0 & 1 & 0 & 0 \\ 0 & 2 & 1
          & 0 & 1 & 0 \\ 0 & 0 & 3 & 7 & 0 & 2 \\ 0 & 0 & 0 & 4 & 9 &
          0 \\ 0 & 0 & 0 & 0 & 9 & 1 \\
        \end{bmatrix}      
        \]}
    \end{column}

  \pause \small{ \[
    \begin{array}{lccccccccccccccc}
      \ve{A}\ values: & 2 & 8 & 4 & 5 & 1 & 2 & 1 & 1 & 3 & 7 & 2 & 4
      & 9 & 9 & 1 \\ column: & 1 & 3 & 1 & 2 & 4 & 2 & 3 & 5 & 3 & 4 &
      6 & 4 & 5 & 5 & 6 \\ row\ start: & 1 & 3 & 6 & 9 & 12 & 14 & 16
      \\
    \end{array}
    \]}

  \pause
  \begin{itemize}
  \item $m$ Krylov iterations require $(m+1)$ subspace vectors
  \item Storage requirement is $\big[(m+1) N\big]$ for $\ve{x} \in
    \mathbb{R}^{N \times N}$
  \end{itemize}

  \pause
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {If we need 10 Krylov iterations to converge...}
    \begin{itemize}
    \item 66 elements required for subspace vectors
    \item 72 total elements required for Jacobian and probability matrix
    \end{itemize}
  \end{beamerboxesrounded}
